\bf\mftitle

### 2011-03-03 doc/jInfer/paper/paper.tex
Documentation: "paper" - review round 1.

\section*{Short architecture overview}
\begin{figure}
	\centering\includegraphics[width=\myscale\columnwidth]{inference_process}
	\caption{High-level view of the inference process} \label{inference_process}
\end{figure}
We divide the inference process into three steps (as illustrated on fig. \ref{inference_process}):
\begin{enumerate}
	\item Import of various formats (XML/XSD/DTD/XPath/\ldots) into an internal representation - \emph{Initial Grammar}.
	\item Simplification (generalization) of \emph{Initial Grammar} with optional help of user interaction.
	\item Export into various schema formats (XSD/DTD/\ldots).
\end{enumerate}	
Our internal representation is a regular grammar, i.e. a list of rules.
We represent the right side of grammar rule as a regular expression with our own set of classes to represent regexps.

@@

\begin{figure}
	\centering\includegraphics[width=\myscale\columnwidth]{ruledisplayer}
	\caption{Visualization of grammar rule} \label{ruledisplayer}
\end{figure}
We provide visualization of rules itself, which is useful for methods not using automatons at all.
Algorithm can use API to display rules to user in progress (see fig. \ref{ruledisplayer}).

\section*{Extensibility}
Developers willing to experiment with their own algorithms have to take care only of \jmodule{Simplifier} module.
We provide sample one, which is divided into submodules again, and again and again.
One can find exact place he wants to replace functionality, and replace only that part of simplification chain.
It then remains configurable whether to use his new experimental module or the old one, from GUI.
He doesn't need to code into jInfer sources, keeping his software installed in same NB instance as jInfer is enough.

As we implement \cite{ahonen}, we provide good support for merging state algorithms - automaton visualization for example.
We divide merging of automaton states in two tasks: merging strategy and merging criterion.
Former one is responsible of deciding whether to merge two states or not.
It has to decide where to stop merging and work is done, it uses merging criterion to test two states equivalency and can measure automaton as it wants to decide whether to continue or stop process.
Merging criterion (we implement $k,h$-context) can be replaced and plugged into merging strategy. One can implement for example $s,k$-strings (\cite{Garofalakis:2000:XSE:335191.335409}).

We provide replacable parts in automaton to regular expression conversion problem, so one can replace our state removal method (with simple weighted heuristic).

@@

\section*{Open issues}
There are still many issues to be worked on.
First is set of supported input/output formats.
On input, one can imagine XML Queries, on output Relax NG or Schematron. 
There is room for implementing schema visualization tool into jInfer, to enable user-aided schema refinement (to approach \cite{Bex:2008:SSI:1376616.1376750}).

For now, we ignore content models of attribute and text node values, but there is interface available to be implemented.
So simple data type detection is waiting for its imlementor.

Many algorithms published wait to be implemented as jInfer modules, maybe some of them will be tried in study of our master theses.

\section*{Conclusion}
Framework presented aims to be the tool used to obtain schemas from XML/XSD/DTD files (and more inputs after extending).
While maintaining distributable bundle for users, with best of module selections and algorithms implemented so far, we predict it will be used in future studies of new algorithms as a test environment.
Best of them will probably migrate into jInfer source tree to aid common users to solve their document set quickly enough, to even bother with schema generation at all.

Weak side of the solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user and
lack of content model inference.
Goal was to produce stable framework for future development and we hope to negate this issues by implementing master theses in this environment, providing it with bundle of different algorithms, solving different input and output formats and infering more concise schemas.

@@

\noindent \begin{tabular}{|l|l|} \hline
Responsible developer & Michal Klempa \\ \hline
Required tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.RuleDisplayer} \\ \hline
Provided tokens       & \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference.Simplifier} \\ \hline
Module dependencies   & AutoEditor, Base, JUNG, Lookup \\ \hline
Public packages       & none \\ \hline
\end{tabular}

%#### INTRO ############################################
\section{Introduction}
In \jmodule{TwoStepSimplifier} we provide the user with a complex solution to grammar simplification problem.
While this is the only simplifier bundled with jInfer, its strong modularity ensures that it can be extended easily. 
For most of the points of extension we indeed provide multiple implementations. 
These vary from trivial proof-of-concept to for example $k,h$-context method for grammar rules inference (see \cite{ahonen}). 
Strong emphasis is put on generic automata, therefore it is easy to plug-in any automaton merging state algorithm.
In this document, we will list all modules and submodules of the simplifier and describe them briefly.

%#### FACTORY ############################################
\section{Factory pattern usage} \label{section_factory}
In \jmodule{TwoStepSimplifier} we employ a factory pattern (as described in \cite[section 4.2]{archdoc}) to divide the module into submodules.
Since service providing classes are being kept as singletons in the NB platform, we use them as factories. For example:
\begin{verbatim}
@ServiceProvider(service = AutomatonSimplifierFactory.class)
public interface AutomatonSimplifierFactory extends 
                     NamedModule, Capabilities, UserModuleDescription {
  <T> AutomatonSimplifier<T> create();
}
\end{verbatim}
The real submodule interface (called familiarly \emph{worker} interface) is the \code{AutomatonSimplifier<T>} interface (an instance of which is returned by the factory).
For now, it is not important what the worker interface looks like, lets just examine the factory.

In \jmodule{Two\.Step\.Sim\.pli\.fier} we design service providing factory interfaces 
so that they extend \code{Na\.med\.Mo\.du\.le}, \code{Ca\.pa\.bi\.li\.ties} and \code{UserModuleDescription}.
Therefore, the modules must implement following methods:
\begin{verbatim}
String getName();
String getDisplayName();
String getModuleDescription();
List<String> getCapabilities();
String getUserModuleDescription();
\end{verbatim}
Nothing non-standard apart from \code{getUserModuleDescription}.
It returns user-friendly description of the module which is then displayed in properties panels.

The \jmodule{AutomatonSimplifier} works with \code{Automaton} instances, which are generic;
but automaton simplification does not depend on any particular type (of the generic symbol) for automaton transitions, 
so interface \code{AutomatonSimplifier<T>} is generic too.
Factory interface deals with this by defining the \code{create()} method also as generic.

Our usage of this factory pattern follows the routine:
\begin{verbatim}
Properties p = RunningProject.getActiveProjectProps(getName());

AutomatonSimplifierFactory f = ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
   p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER));

AutomatonSimplifier<AbstractStructuralNode> autSmp = f.<AbstractStructuralNode>create();
...
result= autSmp.simplify(something_to_process);
\end{verbatim}
If a module has submodules, we implement lookups for submodule implementations in our own factory \code{create()} method.
Worker class receives factories of all of its submodules as a constructor parameters.

Lets look at \code{AutomatonMergingStateFactory}. This is the factory of a module, which has \jmodule{AutomatonSimplifier} as a submodule.
Its create method (shortened): 
\begin{verbatim}
@Override
public ClusterProcessor<AbstractStructuralNode> create() {
  LOG.debug("Creating new ClusterProcessorAutomatonMergingState.");
  return new AutomatonMergingState(getAutomatonSimplifierFactory(),
                                   getRegexpAutomatonSimplifierFactory());
}
\end{verbatim}
Methods \code{getAutomatonSimplifierFactory} and \code{getRegexpAutomatonSimplifierFactory} are analogical. Here is the former:
\begin{verbatim}
private AutomatonSimplifierFactory getAutomatonSimplifierFactory() {
  Properties p = RunningProject.getActiveProjectProps(getName());

  return ModuleSelectionHelper.lookupImpl(AutomatonSimplifierFactory.class,
            p.getProperty(PROPERTIES_AUTOMATON_SIMPLIFIER,
                          PROPERTIES_AUTOMATON_SIMPLIFIER_DEFAULT));
}
\end{verbatim}
Cluster processor \code{AutomatonMergingState} then receives factories 
of \jmodule{AutomatonSimplifier} and \jmodule{Reg\.exp\.Au\.to\.ma\.ton\.Simplifier} submodules in its constructor.
Cluster processor then may create as many instances of submodule classes as it needs (it may be simplifying more than one automaton).
Thorough this document, we will only mention the worker interface when describing a submodule, 
since all factory interfaces are designed the same way as described previously.

%#### STRUCTURE ############################################
\section{Structure}
\jmodule{TwoStepSimplifier} is implemented in package \code{cz.cuni.mff.ksi.jinfer.twostep}.
Module is divided into two classes: \code{TwoStepSimplifier} (main logic) and \code{TwoStepSimplifierFactory} (lookups, interface to other modules).
The latter is registered as service provider of (and implements the) \code{Simplifier} interface 
(defined in package \code{cz.cuni.mff.ksi.jinfer.base.interfaces.inference}).

Its main method \code{start()} receives Initial Grammar in form of
\begin{verbatim}
List<Element> grammar
\end{verbatim}
Each grammar rule is represented by an instance of \code{Element} class. 
Element itself represents the left side of the rule, and its \code{getSubnodes()} method returns 
regular expression representing the right side of the rule (as described in \cite[3.1.1]{archdoc}, they are all concatenations).
Its second parameter is 
\begin{verbatim}
SimplifierCallback callback
\end{verbatim}
Callback which should be invoked when the simplification is done.
On output, simplifier provides Simplified grammar as a parameter of callback function:
\begin{verbatim}
void finished(List<Element> grammar);
\end{verbatim}
Thus simplifier receives Initial Grammar and returns Simplified Grammar, both represented the same way, as a list of elements.

\code{TwoStepSimplifierFactory.start()} creates new \code{TwoStepSimplifier} class instance, providing three factories of submodules to its constructor.
Then it calls \code{simplify(initialGrammar)} method of this instance and its result is passed as parameter to callback function.
And that's all the magic of \code{TwoStepSimplifierFactory} class.

\section{Modular design}
\jmodule{TwoStepSimplifier} is inspired by \cite{1802522} design. Inference proceeds in two steps:
\begin{enumerate}
	\item Clustering the element instances into clusters of (probably) same elements.
	\item Inferring regular expression for each element from examples of element contents taken from all elements in a cluster.
\end{enumerate}
Clustering is delegated to \jmodule{Clusterer} submodule, and the task of regular expression inference 
for each cluster is delegated to \jmodule{ClusterProcessor} submodule. We will examine both of them.
There is also a third submodule called \jmodule{RegularExpressionCleaner}. 
Its purpose is just to beautify output regular expressions, no inference logic is implemented there. 
Modules are drawn on figure \ref{twostep_modules}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=\myscale]{twostep_modules}
	\caption{Submodules of TwoStep simplifier with their implementations} \label{twostep_modules}
\end{figure}
We provide one \jmodule{Clusterer}, four \jmodule{ClusterProcessor} and four \jmodule{RegularExpressionCleaner} implementations.
Each of those will be explained further in this document.

Method \code{TwoStepSimplifier.simplify()} basically does the following.
\begin{verbatim}
// 1. cluster elements
Clusterer<AbstractStructuralNode> clusterer = clustererFactory.create();
clusterer.addAll(initialGrammar);
clusterer.cluster();

// 2. prepare emtpy simplified grammar
List<Element> simplifiedGrammar= new LinkedList<Element>();

// 3. process rules
ClusterProcessor<AbstractStructuralNode> processor =
             clusterProcessorFactory.create();

for (Cluster<AbstractStructuralNode> cluster : clusterer.getClusters()) {
  AbstractStructuralNode node =  
                          processor.processCluster(clusterer, cluster.getMembers());
  RegularExpressionCleaner<AbstractStructuralNode> cleaner = 
                          regularExpressionCleanerFactory.<AbstractStructuralNode>create();
  // 4. add to rules
  simplifiedGrammar.add(
              new Element(node.getContext(),
              node.getName(),
              node.getMetadata(),
              cleaner.cleanRegularExpression(((Element) node).getSubnodes()),
              attList));
}

return simplifiedGrammar;
\end{verbatim}
First, it creates a clusterer, gives it all the rules and performs the clustering.
Then, an empty list of elements is created as the simplified grammar.
For each input rule in \code{initialGrammar}, submodule \jmodule{ClusterProcessor} is invoked to infer the regular expression of that element.
Finally, regular expression cleaning is done by a submodule 
and new \code{Element} instance is created as a copy of processed node, but with inferred and cleaned regexp.

\paragraph{Sentinel processing}
There are some specialities in processing sentinel elements (see \cite[3.2]{archdoc}).
Sentinels can be only on the right side of rules. They have to be sent to clusterer, 
as someone further in the chain can ask for cluster of a sentinel element - which appears on right side.
However, sentinels have no content, and so they must not be taken into account when inferring attribute properties 
(whether it is required or optional), since sentinels have no attributes at all.

We will now examine the submodules for clustering, processing and cleaning.

%#### STRUCTURE->Clusterer ############################################
\subsection{Clusterer submodule}
Clustering is implemented in \code{cz.cuni.mff.ksi.jinfer.twostep.clustering} package.
One cluster is represented by the \code{Cluster<T>} class with \code{T} as type parameter of clustered items.
This class simply holds a \code{Set} of members of the cluster, one of which is chosen as a \code{representant}.

Any clusterer has to comply with the \code{Clusterer} interface.
Its purpose is to cluster a bunch of elements (rules) on input into a bunch of \code{Cluster} class instances (clusters) on output.
It has methods \code{add()} and \code{addAll()} for adding the input items to be clustered.
Main part is the method \code{cluster()}, which does the clustering itself.

As this may be a time-consuming operation, method throws \code{InterruptedException}.
Implementation should take care of checking whether thread is user interrupted (see \cite[p. 12]{archdoc}).
To be able to respond to calls to the \code{T getRepresentantForItem(T item)} method, 
every clusterer implementation has to keep all its clusters even after the clustering had finished. 
Using this method, given an item the correct cluster representative for this item has to be returned.

If no such cluster exists (item was not added for clustering before), we recommend throwing an exception rather than returning \code{null}.
Missing an item probably indicates error in the algorithm, rather than normal workflow.
One can get all the clusters (\code{List<Cluster<T>{}>}) from clusterer by calling \code{getClusters()} method.
Basic usage of the clusterer is:
\begin{verbatim}
Clusterer <T> c = new MyContextClusterer<T>();
c.addAll(initialGrammar);
c.cluster();
...
c.getClusters();
or
c.getRepresentantForItem(x);
\end{verbatim}

%#### STRUCTURE->Clusterer->WithAttributes ############################################
\subsubsection{ClustererWithAttributes extended interface}
You might have noticed that the \jmodule{Clusterer} interface and \code{Cluster} class are generic.
They may be used as design pattern not only for clustering elements in the inference process.
To address clustering of elements in more detail, we created \code{Cluster\.er\.With\.At\.tri\.bu\.tes<T, S>} interface, which extends \code{Clusterer<T>} interface.
It adds method \code{List\.<Cluster<S>{}> get\.At\.tri\.bu\.teClusters(T representant)}, implying that each representative of
type \code{T} (that is a representative of some main cluster) has some ``attribute'' clusters associated with it.
Attribute clusters are of type \code{S} and can be retrieved by calling \code{getAttributeClusters(x)}.

We provide one implementation of this interface described in \ref{section_Iname}.

%#### STRUCTURE->ClusterProcessor ############################################
\subsection{ClusterProcessor module}
\jmodule{ClusterProcessor} takes rules of one cluster of element and obtains a regular expression representing that set of elements. It returns one element, a rule, with name set to desired name of element in schema (not all elements in a cluster have to have same name, if advanced clustering scheme is used, then processor has to choose the right name for the resulting element) and with subnodes set to regular expression inferred.
It processes attributes of all elements in a cluster to obtain meaningful schema attribute specification. These attributes have to be attached to the resulting element.

Interface of submodule is defined as follows.
\begin{verbatim}
public interface ClusterProcessor<T> {
  T processCluster(
          Clusterer<T> clusterer,
          List<T> rules
          ) throws InterruptedException;
}
\end{verbatim}
Why is the cluster processor given a clusterer instance?
Rules themselves contain information about which elements to process, but clusterer has more information about the topic.
Clusterer can tell you the representative for any arbitrary element from input (not only those elements in rules, but also those that may be on right side of rules). Also, clusterer (if it is with attributes) has information about attributes in each cluster.

We describe each \jmodule{ClusterProcessor} implementation shipped in jInfer in sections \ref{section_PassRepresentant} through \ref{section_AutomatonMergingState}.

%#### STRUCTURE->RegularExpressionCleaner ############################################
\subsection{RegularExpressionCleaner module}
In this last subsection, we examine the \jmodule{RegularExpressionCleaner} interface.
Purpose of this submodule is only to make corrections to regular expression, to make them nicer on output.
For example, it is common that converting automaton to regular expression by state elimination produces nested concatenations such as $(name, (person, id))$.
To convert such expression into $(name, person, id)$, one can implement this interface and connect it to work in chain.
Interface definition is straightforward here.
\begin{verbatim}
public interface RegularExpressionCleaner<T> {
  Regexp<T> cleanRegularExpression(Regexp<T> regexp);
}
\end{verbatim}
Given a regular expression, return cleaned regular expression.
Conversion of regular expression is commonly a recursive task and all our implementations work this way.

%#### PREFERENCES ############################################
\section{Preferences}
All settings provided by \jmodule{TwoStepSimplifier} are project-wide, the preferences panel is in \code{cz.cuni.mff.ksi.jinfer.twostep.properties} package.
It is possible to set the following.
\begin{itemize}
	\item Select \jmodule{Clusterer} submodule implementation from those installed.
	\item Select \jmodule{ClusterProcessor} submodule implementation from those installed.
	\item Select \jmodule{RegularExpressionCleaner} submodule implementation from those installed.
\end{itemize}

%#### DATA FLOW ############################################
\section{Data flow}
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{twostep_flow}
	\caption{TwoStep data flow} \label{twostep_flow}
\end{figure}
Process illustrated in figure \ref{twostep_flow} is as follows.
\begin{enumerate}
	\item Initial Grammar on input, which is sent into \jmodule{Clusterer}.
	\item Each cluster is sent to submodule \jmodule{ClusterProcessor}, which returns regular expression for that cluster.
	\item Regular expression is sent to \jmodule{RegularExpressionCleaner} submodule for cleaning.
	\item Regular expressions representing all the processed clusters are added to the list of simplified grammar rules.
	\item List of simplified rules is returned (Simplified grammar).
\end{enumerate}

%#### SUBMODULES ############################################
\section{Submodule implementations}
%#### SUBMODULES->Iname ############################################
\subsection{Iname (ClusteterWithAttributes)} \label{section_Iname}
We cluster elements in \code{Iname\.<Abstract\.Structural\.Node, Attri\.bu\.te>} class, which takes \code{Abstract\.Structural\.No\.de} classes to cluster as main, and \code{Attribute} classes as attributes.
Clustering is done based on element's \code{getName()} equality (ignoring case).
For each rule (element), it is first clustered by finding the cluster which representative has the same name - or create new cluster with this element as representative.
Then we process right side of this element rule (that are nodes from \code{getSubnodes()}).
Since right side of rule is always concatenation (\jmodule{TwoStep} accepts only simple grammar representation), we simply take \code{.getSubnodes().getTokens()} list and iterate through it.
Each node on right side, is examined in the following way.
\begin{itemize}
  \item If it is \textbf{simple data}, throw it to \code{SimpleDataClusterer} class which is associated with main element cluster. It does nothing more, than holds all \code{SimpleData} instances in one cluster. But in future it may be replaced to cluster simple data somehow, to obtain meaningful content models in schemas.
  \item If it is \textbf{element} and it is tagged as sentinel - search main clusters to find cluster with representative of same \code{getName()}, or create new cluster with this sentinel as representative. From IGG, sentinels may be only on right sides of rules and since each element in schemas has to be defined, there must exist another element with same name, which is not sentinel (and maybe it will come to process in future). So there can't be cluster with only one sentinel element in it. 
  \item Do nothing otherwise, since it is element, it has to have its subnodes defined, and therefore it is element that is proper grammar rule and therefore it has to be somewhere in Initial Grammar, thus it is already processed or is on schedule.
\end{itemize}
Attributes of element are processed through a helper attribute clusterer, which is created for each element cluster.
We have bunch of elements of same name in a cluster and one attribute clusterer associated with this bunch.
This attribute clusterer is given all attributes instances encountered in all elements that are in bunch.
It clusters them by name, case insensitive.
Each main cluster also has \code{SimpleData} clusterer associated with it.
Whole scheme is on figure \ref{clusterer_structure}.
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{clusterer_structure}
	\caption{Iname clusterer structure} \label{clusterer_structure}
\end{figure}

Once again, there are main clusters for elements, for each main cluster there are two helper clusterers - one for attribute clusters and one for simple data clusters.

%#### SUBMODULES->PassRepresentant ############################################
\subsection{PassRepresentant (ClusterProcessor)} \label{section_PassRepresentant}
Trivial ClusterProcessor implementation. For each cluster return its representative as the rule to be in resulting schema.
This has nothing to do with grammar simplification, it is just a proof-of-concept submodule.
Input documents are not valid against this odd grammar.
Do not use this in practice, just read the code to understand the bare minimum needed to implement a submodule.

%#### SUBMODULES->Alternations ############################################
\subsection{Alternations (ClusterProcessor)} \label{section_Alternations}
This processor simply gets all right sides from elements in cluster, puts them in one big list and creates alternation regular expression with this list as children.
That is, it creates one big rule with alternation of every positive example observed.
No generalization is done at all.

%#### SUBMODULES->Trie ############################################
\subsection{Trie (ClusterProcessor)} \label{section_Trie}
This processor takes all rules in a cluster, treats them like strings and builds a prefix tree (a ``trie'') of them. More precisely, it takes the first rule and declares it to be a long branch (concatenation of tokens) in a newly created tree. After that, it adds the remaining rules one by one as branches like this: as long as it can follow an existing branch, it follows it. As soon as the newly added branch starts to differ, it ``branches off'' (creates an alternation at that point) the existing tree and hangs the rest of the newly added rule there. Repeating this process creates a prefix tree describing all the rules in the cluster.
The tree is the final regular expression.

%#### SUBMODULES->AutomatonMergingState ############################################
\subsection{AutomatonMergingState (ClusterProcessor)} \label{section_AutomatonMergingState}
\begin{figure}[tpb]
	\centering\includegraphics[scale=\myscale]{automaton_merging_state_modules}
	\caption{Submodules of AutomatonMergingState cluster processor} \label{automaton_merging_state_modules}
\end{figure}
\begin{figure}[tpb]
	\centering\includegraphics[scale=1]{automaton_merging_state_flow}
	\caption{Data flow of AutomatonMergingState cluster processor} \label{automaton_merging_state_flow}
\end{figure}
\jmodule{AutomatonMergingState} is an implementation of merging state algorithm on nondeterministic finite automaton (see package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate}).
\subsubsection{Structure} \label{section_AutomatonMergingState_structure}
Abridged code of cluster processor operation follows.
\begin{verbatim}
// 1. Construct PTA
Automaton<AbstractStructuralNode> automaton = new Automaton<AbstractStructuralNode>(true);

// Take each rule in cluster and pass right side to automaton to create PTA
for (AbstractStructuralNode instance : rules) {
  Element element = (Element) instance;
  Regexp<AbstractStructuralNode> rightSide = element.getSubnodes();

  List<AbstractStructuralNode> rightSideTokens = rightSide.getTokens();

  List<AbstractStructuralNode> symbolString = new LinkedList<AbstractStructuralNode>();
  for (AbstractStructuralNode token : rightSideTokens) {
    symbolString.add(clusterer.getRepresentantForItem(token));
  }
  automaton.buildPTAOnSymbol(symbolString);
}

// 2. Simplify automaton by merging states using automatonSimplifier
Automaton<AbstractStructuralNode> simplifiedAutomaton =
    automatonSimplifier.simplify(automaton, elementSymbolToString);

// 3. Convert Automaton<AbstractStructuralNode> to RegexpAutomaton<AbstractStructuralNode>
RegexpAutomaton<AbstractStructuralNode> regexpAutomaton =
    new RegexpAutomaton<AbstractStructuralNode>(simplifiedAutomaton);

// 4. Call regexpAutomatonSimplifier to obtain regular expression from regexpAutomaton
Regexp<AbstractStructuralNode> regexp =
    regexpAutomatonSimplifier.simplify(regexpAutomaton, regexpAbstractToString);

// 5. Return element with regexp
return new Element(
      new ArrayList<String>(),
      rules.get(0).getName(),
      new HashMap<String, Object>(),
      regexp,
      new ArrayList<Attribute>()
      );
\end{verbatim}
The whole submodule structure is in figure \ref{automaton_merging_state_modules}.

\subsubsection{Data flow}
\jmodule{AutomatonMergingState} works in following steps.
\begin{enumerate}
	\item Create prefix-tree automaton (PTA) from positive examples - right sides of rules.
Call its submodule \jmodule{Au\.to\.ma\.ton\.Simpli\.fier} to modify PTA to some generalized automaton by merging states.
	\item Simplified automaton is then converted to an instance of \code{RegexpAutomaton} (see \ref{section_RegexpAutomaton}) by using clone constructor.
Regexp automaton is automaton with regular expression as symbol on transitions.
In automata theory, such automaton is called extended NFA.
Clone constructing is done by converting each symbol in source automaton to regexp token with that symbol as \code{content}.
	\item \code{RegexpAutomaton} is then passed into second submodule called \jmodule{RegexpAutomatonSimplifier}.
Its job is to derive regular expression from automaton, such that automaton and regular expression represent the same language.
\end{enumerate}
The data flow is depicted in figure \ref{automaton_merging_state_flow}.

\subsubsection{Data structures - Automaton} \label{section_Automaton}
\begin{figure}[tpb]
	\centering\includegraphics[scale=\myscale]{pta}
	\caption{PTA created from input strings: $aa, aaaaa, abaa, bab, baa$} \label{pta}
\end{figure}
Class \code{Automaton} is shortly described in \cite{archdoc}.
Let's take a look at method \code{build\.PTA\.On\.Sym\.bol} here.
Given \code{List<T> symbolString} it traverses automaton - comparing transition symbols with symbols in string.
It follows transitions in automaton until first difference with string symbol is found (like in prefix trees).
On different symbol, it creates new branch and all states and transitions on that branch.
Last state visited gets its \code{finalCount} value incremented, as one more input string ended in this state.
The code is as follows.
\begin{verbatim}
public void buildPTAOnSymbol(List<T> symbolString) {
  State<T> xState= this.getInitialState();
  for (T symbol : symbolString) {
    Step<T> xStep= this.getOutStepOnSymbol(xState, symbol);
    if (xStep != null) {
      xStep.incUseCount();
      xState= xStep.getDestination();
    } else {
      State<T> newState= this.createNewState();
      Step<T> newStep= this.createNewStep(symbol, xState, newState);
      assert newStep.getDestination().equals(newState);
      xState= newStep.getDestination();
    }
  }
  xState.incFinalCount();
}
\end{verbatim}
Starting from initial state, traverse the automaton by asking for \code{getOutStepOnSymbol()}.
If there is such step (transition), follow it to its destination state and move the position in \code{symbolString}.
If there is no such step, create one and create also the destination state, then follow this newly created step to new state.

As you can see, it is important that class \code{T} has proper implementation of \code{.equals()} method, as symbols are tested for equality by calling this method on them.
By the way, this is why we cluster all simple data and all sentinels on right sides of rules.	
While adding a concatenation to the automaton, instead of the instances themselves their \emph{representatives} are used. See \ref{section_AutomatonMergingState_structure} for further reference.
This assures that comparisons in automaton are done using Java \code{Object.equals()} reference comparison, but for each cluster member, its representative is only present in automaton.
That results in cluster-equal behaviour.
Example of prefix-tree automaton is in figure \ref{pta}.

\subsubsection{Data structures - RegexpAutomaton} \label{section_RegexpAutomaton}
Class \code{RegexpAutomaton} extends class \code{Automaton} with generic type \code{T} set to \code{Regexp<T>}.
That is, you have to instantiate it with same symbol type (e.g. \code{AbstractStructuralNode}) as automaton.
Implementation is in package \code{cz.cuni.mff.ksi.jinfer.twostep.processing.automatonmergingstate.regexping}. This is the same package where \jmodule{RegexpAutomatonSimplifier} interface is defined.
Implementation is nothing special - there is a copy constructor which creates regular expression automaton from ordinary automaton by enclosing each symbol on transition of original automaton in a \code{Regexp.TOKEN}.

\subsubsection{AutomatonSimplifier module}
\jmodule{AutomatonSimplifier} module has its worker interface defined as following.
\begin{verbatim}
public interface AutomatonSimplifier<T> {
  Automaton<T> simplify(Automaton<T> inputAutomaton, 
                        SymbolToString<T> symbolToString) throws InterruptedException;
  Automaton<T> simplify(Automaton<T> inputAutomaton, 
                        SymbolToString<T> symbolToString,
                        String elementName) throws InterruptedException;
}
\end{verbatim}
Given input automaton it returns automaton, there is nothing magical to it.
Interface is defined as generic, implementations don't have to be. However, there is no reason not to make them generic.
When implementation needs to present an automaton to user, it needs string representation of symbols to display on automaton transitions.
For this reason, the second parameter, \code{symbolToString} (implementation of \code{SymbolToString} interaface) is given.
It is responsible for converting symbol to string.
Overloaded version is to inform the user about name of processed element when presenting automaton in \jmodule{AutoEditor}.
This version of interface will probably change in future.

This generality means that simplifier can not only be used to infer rules for elements, but possibly also to infer patterns strings for content model of attributes using the same algorithms.
The actual implementation of this still waits for its developer to come.
Implementations of \jmodule{AutomatonSimplifier} are discussed in \ref{section_Greedy} and \ref{section_UserInteractive}.

\subsubsection{RegexpAutomatonSimplifier submodule}
Purpose of \jmodule{RegexpAutomatonSimplifier} is to obtain a regular expression from a given automaton.
Interface for doing is is thus simple.
\begin{verbatim}
public interface RegexpAutomatonSimplifier<T> {
  Regexp<T> simplify(RegexpAutomaton<T> inputAutomaton, 
                     SymbolToString<Regexp<T>> symbolToString) throws InterruptedException;
}
\end{verbatim}
Once again we encounter \code{symbolToString} with the same purpose as before, if submodule would need to present the automaton to user, it has to be able to convert symbols of (at runtime) unknown type to strings.
Our implementation of \jmodule{RegexpAutomatonSimplifier} is discussed in \ref{section_StateRemoval}

\subsubsection{MergeConditionTester submodule}
\jmodule{AutomatonMergingState} has one more submodule, the \jmodule{MergeConditionTester}, which is not called directly by \jmodule{Au\.to\.ma\.ton\.Merg\.ing\.Sta\.te}.
It is at disposal of implementations of \code{AutomatonSimplifier} interface for testing, whether two states in automaton are equivalent and should be merged into one state.

One can implement ACO heuristics or MDL principle heuristics as \jmodule{AutomatonSimplifier} submodule, and can use any of \jmodule{MergeConditionTesters} provided.

\subsection{NestedConcatention (RegularExpressionCleaner)} \label{section_NestedConcatention}
It converts nested concatenations into flat ones.
Example: $(name, (person, id))$ is converted to $(name, person, id)$.

\subsection{Null (RegularExpressionCleaner)} \label{section_Null}
This cleaner just returns regexp it receives and does nothing.
Plug this one into chain, if you want to omit this step in inference.

\subsection{EmptyChildren (RegularExpressionCleaner)} \label{section_EmptyChildren}
Wipes out regular expressions of type: concatenation, alternation, permutation, which have empty \code{children} member.
If for example an inference method produces unnecessary empty regexps in regexp tree such as
$$
(name, (), (person, id))
$$
There is one empty concatenation/alternation/permutation with no sense.
This is wiped out by this cleaner to produce 
$$
(name, (person, id))
$$

\subsection{Chained (RegularExpressionCleaner)} \label{section_Chained}
Allows the user to chain more cleaners with output from first one plugged as input into second one and so one.
Thus regular expression from inference can pass through \jmodule{EmptyChildren} and then through \jmodule{NestedConcatenation} cleaners and then it is returned.

\subsection{Greedy (AutomatonSimplifier)} \label{section_Greedy}
We implement greedy strategy of automaton simplifying in class \code{Greedy}.
It simply asks given \code{MergeConditionTester} if it can merge any of automaton states and merges states until there are no states to be merged.
\begin{verbatim}
while (there exist pair of states that are equivalent) {
  merge them
}
\end{verbatim}
We provide one \jmodule{MergeConditionTester} implementation, see \ref{section_KHContext}.

>>>>

\section*{Short architecture overview}
\begin{figure}
	\centering\includegraphics[width=\myscale\columnwidth]{inference_process}
	\caption{High-level view of the inference process} \label{inference_process}
\end{figure}
We divide the inference process into three steps (as illustrated in fig. \ref{inference_process}):
\begin{enumerate}
	\item Import of various formats (XML/XSD/DTD/XPath/\ldots) into an internal representation - \emph{Initial Grammar}.
	\item Simplification (generalization) of \emph{Initial Grammar} with optional help of user interaction.
	\item Export into various schema formats (XSD/DTD/\ldots).
\end{enumerate}	
Our internal representation is a regular grammar, i.e. a list of rules.
We represent the right side of grammar rule as a regular expression with our own set of classes to represent regexps.

@@

\begin{figure}
	\centering\includegraphics[width=\myscale\columnwidth]{ruledisplayer}
	\caption{Visualization of a grammar rule} \label{ruledisplayer}
\end{figure}
We provide visualization of rules themselves, which is useful for methods not using automata at all.
Algorithm can use API to display rules to user in progress (see fig. \ref{ruledisplayer}).

\section*{Extensibility}
Developers willing to experiment with their own algorithms need to take care only of \jmodule{Simplifier} module.
We provide a sample one, which is recursively divided into submodules.
One can find the exact place to replace functionality he wants, and replace only that part of the simplification chain.
It is still optional whether to use this new experimental module or the old, bundled one.
He doesn't need to code into jInfer sources, keeping his software installed in same NB instance as jInfer is enough. %TODO

As we implement \cite{ahonen}, good support for merging state algorithms is provided - automaton visualization for example.
We divide automaton state merging in two tasks: merging strategy and merging criterion.
Former one is responsible of deciding whether to merge two states or not.
It has to decide where to stop merging and work is done, it uses merging criterion to test two states equivalency and can measure automaton as it wants to decide whether to continue or stop process.
Merging criterion (we implement $k,h$-context) can be replaced and plugged into the merging strategy. One can implement for example $s,k$-strings (\cite{Garofalakis:2000:XSE:335191.335409}).

We provide replacable parts in automaton-to-regexp conversion problem, so one can replace our state removal method (with simple weighted heuristic).

@@

\section*{Open issues}
There are still many issues to be worked on.
First one is the set of supported input/output formats.
On input, one can imagine support for XML Queries, on output export to languages like Relax NG or Schematron. 
There is room for implementing schema visualization tool into jInfer, to enable user-aided schema refinement (to approach \cite{Bex:2008:SSI:1376616.1376750}).

For now we ignore content models of attribute and text node values, but there is an interface waiting to be implemented.
Thus, simple data type detection is waiting for its implementor.

Many algorithms published wait to be implemented as jInfer modules, perhaps some of them will be implemented in the course of our master theses.

\section*{Conclusion}
The framework presented in this paper aims to be the tool used to obtain schemas from XML/XSD/DTD files (and more after extension).
While maintaining a distributable bundle for users, with the best modules and algorithms implemented so far, we predict it will be used in future studies of new algorithms as a test environment.
Best of them will probably merged into jInfer source tree to aid common users to solve their schema inference problems quickly and easily.

Weak side of our solution is the lack of good quality automatic or semi-automatic algorithms to be used by common user, as well as the lack of content model inference.
Our goal was to produce a stable framework for future development and we hope to negate these issues by implementing our master theses in this environment, providing it with a bundle of different algorithms, solving different input and output formats and infering more concise schemas.

@@
